---
title: "Homework 3"
author: "Keyanna Davis"
date: "10/5/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(knitr)
library(p8105.datasets)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

#Problem 0 
This “problem” focuses on structure of your submission, especially the use git 
and GitHub for reproducibility, R Projects to organize your work, R Markdown 
to write reproducible reports, relative paths to load data from local files, 
and reasonable naming structures for your files.
  
To that end:


* create a public GitHub repo + local R Project; we suggest naming this repo / 
directory p8105_hw3_YOURUNI (e.g.  p8105_hw3_ajg2202 for Jeff), but that’s not 
required
* create a single .Rmd file named p8105_hw3_YOURUNI.Rmd that renders to 
github_document


Your solutions to Problems 1, 2, and 3 should be implemented in your .Rmd file, 
and your git commit history should reflect the process you used to solve these 
Problems.


For this Problem, we will assess adherence to the instructions above regarding 
repo structure, git commit history, and whether we are able to knit your .Rmd 
to ensure that your work is reproducible. Adherence to appropriate styling and 
clarity of code will be assessed in Problems 1+ using the homework style rubric.


This homework includes figures; the readability of your embedded plots (e.g. 
font sizes, axis labels, titles) will be assessed in Problems 1+.


#Problem 1

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):

```{r}
data("instacart") 
instacart

```

The instacart dataset has `r nrow(instacart)` rows and `r ncol(instacart)` columns in the data. All variables are class `numeric` or `character`. Each row contains information about an item that was ordered. The variables include the order information such as order id, reorders, order day of week, etc. The variables also give information about the item such as product name, product id, aisle, department, etc. There are `r n_distinct(pull(instacart,department))` different departments and `r n_distinct(pull(instacart, product_name))` different products from these departments in this dataset. 


####Problem 1.1
* How many aisles are there, and which aisles are the most items ordered from?
```{r}
instacart %>% 
  count(aisle) %>% 
   arrange(desc(n)) %>% 
  knitr::kable()
```


There are `r n_distinct(pull(instacart, aisle))` aisle in this dataset. `Fresh vegatables`,  `fresh fruits`, and `package vegatables fruits` has the most items ordered from. 


####Problem 1.2
* Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
 instacart %>% 
   group_by(aisle) %>% 
   summarize(n = n()) %>% 
   filter(n > 10000) %>% 
   ggplot(aes( x = n, y = aisle, color = aisle)) +
  geom_point() +
  labs(
    title = "Aisles with more than 10000 Orders",
    x = "total number of items ordered in each aisle",
    y = "aisles"
  ) +
  theme(legend.position = "none")
```

According to the plot above the most orders were from `fresh vegatables` and `fresh fruits` about around `150000`. The plot also shows there were alot of aisles with the total amount of items orders ranging from `10000` and `20000`. 


####Problem 1.3
* Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.



In order for me to make this table I will make a subset from the dataset filtering out the `3` particular aisles separately combine the `3` datasets to make one table. 
```{r}
BI = instacart %>% 
     filter(aisle%in% c("baking ingredients") ) %>% 
   group_by(aisle, product_name) %>% 
   summarize(n=n()) %>% 
     arrange(desc(n)) %>% 
     head(3)

DC = instacart %>% 
     filter(aisle%in% c("dog food care") ) %>% 
   group_by(aisle, product_name) %>% 
   summarize(n=n()) %>% 
     arrange(desc(n)) %>% 
     head(3)  

PVF = instacart %>% 
     filter(aisle%in% c("packaged vegetables fruits") ) %>% 
   group_by(aisle, product_name) %>% 
   summarize(n=n()) %>% 
     arrange(desc(n)) %>% 
     head(3)  

rbind(BI, DC, PVF) %>%
  arrange(desc(n)) %>% 
  knitr::kable()

```


The above table shows that the `3` most popular items from the aisles `packaged vegatables fruits`, `baking ingredients` and `dog food care`. `Organic Baby Spinach` is the most popular item from `packaged vegatables fruits` and the most popular of the `3` aisles. `Pure Baking Soda` is the second most popular item of the `baking ingredients` aisle. `Small Dog Biscuits` is the third most popular item of the `dog food care` aisle. 


####Problem 1.4
* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
apples_and_icecream = instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(average_hour = mean(order_hour_of_day)) %>%
  select(product_name, order_dow, average_hour) %>% 
  mutate(
    week_day = recode(
      order_dow,
      '0' = "Sunday",
      '1' = "Monday",
      '2' = "Tuesday",
      '3' = "Wednesday",
      '4' = "Thursday",
      '5' = "Friday",
      '6' = "Saturday"
    )
  ) %>%
  select(product_name, week_day, average_hour) %>% 
  pivot_wider(
    names_from = "week_day",
    values_from = "average_hour") %>% view()
apples_and_icecream %>% 
  knitr::kable()

```

The above table shows the average hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. On `Sunday` the average hour of the day is about the same for `Coffee Ice Cream` and `Pink Lady Apples` around `13:00`. Also on `Friday` the average hour of the day is about the same for both `Coffee Ice Cream` and `Pink Lady Apples` around `12:00`. 



#Problem 2
* do some data cleaning
* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels from “Excellent” to “Poor”

```{r}
data("bfrss_smart2010") 

brfss = brfss_smart2010 %>% 
rename("state" = "Locationabbr",
       "county" = "Locationdesc"
       ) %>% 
  mutate(
    county = str_sub(county, 5)
  ) %>% 
     mutate(
     Response = factor(Response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))
   ) %>%
   arrange(desc(Response)) %>% 
  filter(Topic=="Overall Health") %>% 
  select(-Question,-(Confidence_limit_Low:GeoLocation)) %>% 
  drop_na()
 
```


####Problem 2.1
* In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
brfss %>% 
  filter(Year %in% "2002") %>% 
  group_by(state) %>% 
  summarize(n_location = n_distinct(county)) %>% 
  filter(n_location >= 7) %>% 
  knitr::kable()

brfss %>% 
  filter(Year %in% "2010") %>% 
  group_by(state) %>% 
  summarize(n_location = n_distinct(county)) %>% 
  filter(n_location >= 7) %>%
  knitr::kable(caption = "States with 7 or more Locations")
  
```

In `2002` the states that had 7 or more locations were `Connecticut, FLorida, Massachusetts, North Carolina , New Jersey, and Pennsylvania`. In `2010` the states that had 7 or more locations were `California, Colorado, Florida, Massachusetts, Maryland, North Carolina, Nebraska, New Jersey, New York, Ohio, Pennsylvania, South Carolina, Texas and Washington`. 


####Problem 2.2
* Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help)

```{r}
brfss %>% 
  filter(Response == "Excellent") %>% 
  select(Response, Year, state, Data_value) %>% 
  group_by(Response, Year, state) %>% 
  summarise(avg_data_value = mean(Data_value)) %>% 
  ggplot(aes(x = Year, y= avg_data_value, group=state, color = state))+
  geom_line()
```

The plot shown above indicates that the in `2005` `West Vigrinia` has the lowest `averaged data value` among the rest of the states throughtout the years. In `2003` `Connecticut` has the highest `averaged data value` among the rest of the states throughout the years.  



####Problem 2.3
* Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss %>%
  filter(Year %in% c("2006", "2010"), state == "NY") %>% 
  ggplot(aes(x=Response, y=Data_value, color=Response)) +
  geom_boxplot(alpha = .5) +
  facet_grid(~Year)  + 
  viridis::scale_fill_viridis(discrete = TRUE)
```



The above diagram shows a two panel boxplots distributions of data value for responses `Excellent, Very good, Good, Fair, and Poor` for the years `2006` and `2010`. For `Excellent Response` the distribution of data value is skewed to the right for both years `2010` and `2006`. For year `2006` `Good Response` and `Fair Response` distributions of data value are both skewed to the right. In `2006` and `2010` `Very Good Response` distribution data value was skewed to the left. According to the plot above the in `2006` the `Poor Response` the data value was normally distributed but in `2010` it was skewed to the right. In `2010` the `Fair Response` was the data value was normally distributed. Also in `2010` the `Good Response` data value was skewed to the left. 





#Problem 3

Accelerometers have become an appealing alternative to self-report techniques for studying physical activity in observational studies and clinical trials, largely because of their relative objectivity. During observation periods, the devices measure “activity counts” in a short period; one-minute intervals are common. Because accelerometers can be worn comfortably and unobtrusively, they produce around-the-clock observations.

This problem uses five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). The data can be downloaded here. **In this spreadsheet, variables `activity.* ` are the activity counts for each minute of a 24-hour day starting at midnight.**

*Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
```{r}
accel_data = read_csv("~/Desktop/p8105_hw3_kd2640/data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minutes",
    names_prefix = "activity_",
    values_to= "counts"
  )  %>% 
  mutate(
    day_of_week = if_else(day=="Saturday"| day=="Sunday", "weekend", "weekday")
  ) %>% 
  mutate(
   minutes = as.numeric(minutes)
  )
nrow((accel_data))
```

The accelerometer dataset has `r nrow(accel_data)` observations and `r ncol(accel_data)` variables including `week`, `day_id`, `day`, `minutes`, `counts`, and `day_of_week`. 


*Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel_data2 =
  accel_data %>% 
  select(week, minutes, day, counts) %>% 
  group_by(week, day) %>% 
  summarize(total_activity = sum(counts)) 

  accel_data2 %>% 
  knitr::kable()

  
 accel_data2 %>%  
   ggplot(aes(x = week, y = total_activity, color = day)) +
  geom_line() +
  geom_point()


  
```

The above graph shows a `spaghetti plot`. On `Wednesday` the activity is steady throughout the weeks. I don't see any trends on the other days of the week it seems the activity is up and down througout the `5` weeks. 



####Problem 3.3
* Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.


```{r}
accel_data %>% 
  ggplot(aes(x = minutes, y = counts, color = day, group = day_id)) +
  geom_line()
```

