---
title: "Homework 3"
author: "Keyanna Davis"
date: "10/5/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(knitr)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


#Problem 1

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):

```{r}
load("~/Downloads/instacart.rda")
 instacart %>% 
  count(aisle) %>% 
   arrange(desc(n))

```

The instacart dataset has `r nrow(instacart)` rows and `r ncol(instacart)` columns in the data. Each row contains information about an item that was ordered. The variables include the order information such as order id, reorders, order day of week, etc. The variables also give information about the item such as product name, product id, aisle, department, etc. 




```{r}
 instacart %>% 
   group_by(aisle) %>% 
   summarize(n = n()) %>% 
   filter(n > 10000) %>% 
   ggplot(aes( x = n, y = aisle, color = aisle)) +
  geom_point()
```


```{r}
BI = instacart %>% 
     filter(aisle%in% c("baking ingredients") ) %>% 
   group_by(aisle, product_name) %>% 
   summarize(n=n()) %>% 
     arrange(desc(n)) %>% 
     head(3)

DC = instacart %>% 
     filter(aisle%in% c("dog food care") ) %>% 
   group_by(aisle, product_name) %>% 
   summarize(n=n()) %>% 
     arrange(desc(n)) %>% 
     head(3)  

PVF = instacart %>% 
     filter(aisle%in% c("packaged vegetables fruits") ) %>% 
   group_by(aisle, product_name) %>% 
   summarize(n=n()) %>% 
     arrange(desc(n)) %>% 
     head(3)  

rbind(BI, DC, PVF) %>% 
  knitr::kable()
```


```{r}
apples_and_icecream = instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(average_hour = mean(order_hour_of_day)) %>%
  select(product_name, order_dow, average_hour) %>% 
  mutate(
    week_day = recode(
      order_dow,
      '0' = "Sunday",
      '1' = "Monday",
      '2' = "Tuesday",
      '3' = "Wednesday",
      '4' = "Thursday",
      '5' = "Friday",
      '6' = "Saturday"
    )
  ) %>%
  select(product_name, week_day, average_hour) %>% 
  pivot_wider(
    names_from = "week_day",
    values_from = "average_hour") 
apples_and_icecream %>% 
  knitr::kable()
```






#Problem 2
* do some data cleaning
* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels from “Excellent” to “Poor”

```{r}
load("~/Downloads/brfss_smart2010.rda")

brfss = brfss_smart2010 %>% 
rename("state" = "Locationabbr",
       "county" = "Locationdesc"
       ) %>% 
  mutate(
    county = str_sub(county, 5)
  ) %>% 
  filter(Topic=="Overall Health", Response %in% c("Excellent", "Very Good", "Good", "Fair", "Poor"))%>% 
  arrange(desc(Response)) %>% 
  select(-Question,-(Confidence_limit_Low:GeoLocation)) 
  
  
```

####Question 2.1
In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
brfss %>% 
  filter(Year %in% "2002") %>% 
  group_by(state) %>% 
  summarize(n_location = n_distinct(county)) %>% 
  filter(n_location >= 7) 

brfss %>% 
  filter(Year %in% "2010") %>% 
  group_by(state) %>% 
  summarize(n_location = n_distinct(county)) %>% 
  filter(n_location >= 7)
  
```

In 2002 the states that had 7 or more locations were Connecticut, FLorida, Massachusetts, North Carolina , New Jersey, and Pennsylvania. In 2010 the states that had 7 or more locations were California, Colorado, Florida, Massachusetts, Maryland, North Carolina, Nebraska, New Jersey, New York, Ohio, Pennsylvania, South Carolina, Texas and Washington. 


####Question 2.2
Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help)

```{r}
brfss %>% 
  filter(Response == "Excellent") %>% 
  select(Response, Year, state, Data_value) %>% 
  group_by(Response, Year, state) %>% 
  summarise(avg_data_value = mean(Data_value)) %>% 
  ggplot(aes(x = avg_data_value, y= state, color = Year))+
  geom_line()
```


####Question 2.3
Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss %>%
  filter(Year %in% c("2006", "2010"), state == "NY") %>% 
  ggplot(aes(x=Data_value, fill=Response)) +
  geom_density(alpha = .5) +
  facet_grid(~Year)  + 
  viridis::scale_fill_viridis(discrete = TRUE)
```








#Problem 3

Accelerometers have become an appealing alternative to self-report techniques for studying physical activity in observational studies and clinical trials, largely because of their relative objectivity. During observation periods, the devices measure “activity counts” in a short period; one-minute intervals are common. Because accelerometers can be worn comfortably and unobtrusively, they produce around-the-clock observations.

```{r}
accel_data = read_csv("~/Desktop/p8105_hw3_kd2640/data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
head() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity",
    values_to= " "
  ) %>% view 
```



